Topics Causal Inference

### 2024-08-16 documents/paper.tex
Add estimated CATE for wgan MSE.

\paragraph*{WA DGP}
I use the third DGP discussed in~\cite{wager2018estimation} as my baseline DGP\@.
Covariates are generated as $X~U([0, 1]^d)$ and potential outcomes are drawn from $Y^{(0,1)} \sim N(\E[Y^{(0,1)}|X], 1)$.
~\cite{wager2018estimation} refer to $d$ as the ``ambient dimension''. Note that in their simulations this is \textit{not} the dimensionality of the CATE function.
Instead, the CATE function is generated by
\begin{equation*}
    \tau(x) = \xi(X_1)\xi(X_2), \qquad \xi(x) = \frac{2}{1 + \exp^{-12(x-1/2)}}.
\end{equation*}
That is, the CATE function only depends on two covariates and the other $d-2$ covariates are pure noise (in terms of CATE estimation).
The propensity score is kept constant at 0.5.

\paragraph*{WGAN Approximation}
Describe how WGAN was performed.

@@

\begin{table}
    \caption{Simulation Results: WA Replication and WGAN}
    \center
    \begin{tabular}{lccccc}
        & Dimension & MSE & SD & Coverage & SD \\
        \toprule
        \multicolumn{6}{l}{\textit{Panel A:\ Original WA DGP 3}} \\
        \input{../bld/wa_replication/tables/wa_replication_dgp3.tex} \\
        \midrule

        \multicolumn{6}{l}{\textit{Panel B:\ High Dimensional --- 4 Covariates}} \\
        \input{../bld/wa_replication/tables/wa_replication_dgp4.tex} \\
        \midrule

        \multicolumn{6}{l}{\textit{Panel C:\ High Dimensional --- 8 Covariates}} \\
        \input{../bld/wa_replication/tables/wa_replication_dgp5.tex} \\
        \midrule

        \multicolumn{6}{l}{\textit{Panel D:\ Placebo}} \\
        \input{../bld/wa_replication/tables/wa_replication_dgp6.tex} \\
        \bottomrule
    \end{tabular}

>>>>

\paragraph*{WA DGP}
I use the third DGP discussed in~\cite{wager2018estimation} as my baseline DGP\@.
Covariates are generated as $X~U([0, 1]^d)$ and potential outcomes are drawn from $Y^{(0,1)} \sim N(\E[Y^{(0,1)}|X], 1)$.
~\cite{wager2018estimation} refer to $d$ as the ``ambient dimension''. Note that in their simulations this is \textit{not} the dimensionality of the CATE function.
Instead, the CATE function is generated by
\begin{equation*}
    \tau_1(x) = \xi(x_1)\xi(x_2), \qquad \xi(x) = \frac{2}{1 + \exp^{-12(x-1/2)}}.
\end{equation*}
That is, the CATE function only depends on two covariates and the other $d-2$ covariates are pure noise (in terms of CATE estimation).
The propensity score is kept constant at 0.5. I present results for various ambient dimensions $d$.

I consider three additional DGPs: The first two DGPs increase the dimension of the CATE function, by increasing the number of arguments the function depends on.
In particular, I use
\begin{equation*}
    \tau_2(x) = \prod_{i=1}^{4}\xi(x_i),
\end{equation*}
as well as
\begin{equation*}
    \tau_3(x) = \prod_{i=1}^{8}\xi(x_i).
\end{equation*}
I keep the rest of the DGP the same. Finally I also consider a \textit{placebo} setting with
\begin{equation*}
    \tau_4(x) = 0.
\end{equation*}
The first two extensions are used to investigate the performance of the causal forest and generic machine learning approach when heterogeneity is high-dimensional.
The placebo extension is used to investigate the danger of over-fitting where machine learning algorithms might be predict a heterogeneous treatment effect when there is actually none.
A similar question is investigated in Figure 1 of~\cite{chernozhukov2023genml}.

\paragraph*{WGAN Approximation}
For generating the WGAN approximation to the DGP I follow the baseline approach described in ds-wgan package example.
In particular, I train the generator and critic using the same hyperparameters.
All fitting is performed twice, once for the distribution of $Y$ conditional on $X, D$ and once for the distribution of $X$ conditional on $D$.
I train the model based on a sample of $N=100,000$ observations with dimension $d=10$ drawn from the baseline WA dgp as well as the extensions with high-dimensional CATEs ($\tau_2, \tau_3$).
I deliberately chose a large sample (at least larger than the $N=10,000$ observations used throughout simulations) to attenuate noise resulting from this original sample draw.
To further address this it would be ideal to increase $N$ further or aggregate the results over different initial sample draws.
Given the training time of these models I found both approaches to be infeasible for the scope of this paper.

One additional difficulty in using WGANs for CATE estimation is to define the target parameter/the "truth" against to which we want to evaluate the estimates.
In the simpler case of ATE estimation, one approach is to draw a large sample from the generator, which we term the population, and to compute the ATE in this population by
\begin{equation*}
    ATE := \frac{1}{N^{pop}} Y_i(1) - Y_i(0).
\end{equation*}
However, a similar approach is not available for the CATE.
While we could compute a similar average conditional on $X$, if there are continuous covariates that would be a bad approximation to the true underlying CATE since we end up with individual treatment effects.
If, on the other hand, we fit a model to the population we know from that even if the WGAN preserved the high-dimensional structure it would not be possible to consistently estimate it.
For the analysis I compute MSE using two approaches: Once using the true CATE function of the WA DGP and once using an estimated CATE based on the generated population using a causal forest approach.
The second approach in particular should result in understated MSE estimates: If the WGAN generator cannot approximate the high-dimensional structure, using an estimated CATE based on a generated population will understate the difficulty of the problem.
The first approach is of course not available in practice since the true CATE is unknown.

@@

\begin{table}
    \caption{Simulation Results: WA Replication and WGAN}
    \center
    \begin{tabular}{lccccccc}
        & Dimension & MSE (True) & SD & MSE (Pop) & SD & Coverage & SD \\
        \toprule
        \multicolumn{8}{l}{\textit{Panel A:\ Original WA DGP 3}} \\
        \input{../bld/wa_replication/tables/wa_replication_dgp3.tex} \\
        \midrule

        \multicolumn{8}{l}{\textit{Panel B:\ High Dimensional --- 4 Covariates}} \\
        \input{../bld/wa_replication/tables/wa_replication_dgp4.tex} \\
        \midrule

        \multicolumn{8}{l}{\textit{Panel C:\ High Dimensional --- 8 Covariates}} \\
        \input{../bld/wa_replication/tables/wa_replication_dgp5.tex} \\
        \midrule

        \multicolumn{8}{l}{\textit{Panel D:\ Placebo}} \\
        \input{../bld/wa_replication/tables/wa_replication_dgp6.tex} \\
        \bottomrule
    \end{tabular}
